\documentclass{anstrans}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{upgreek}
\usepackage[capitalise,nameinlink]{cleveref}

\newcommand{\efr}{Q_\text{fr}}
\newcommand{\enp}{Q_\text{n,p}}
\newcommand{\ened}{Q_\text{n,d}}
\newcommand{\egp}{Q_\text{$\gamma$,p}}
\newcommand{\egd}{Q_\text{$\gamma$,d}}
\newcommand{\eb}{Q_\upbeta}
\newcommand{\enu}{Q_\nu}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Energy Deposition in the OpenMC Monte Carlo Particle Transport Code}
\author{Paul K.~Romano,$\!^{*}$ Andrew E. Johnson,$\!^{\dagger}$ Amanda L.~Lund,$\!^{*}$ and Jingang Liang$^{\ddag}$}

\institute{
$^{*}$Argonne National Laboratory, 9700 S. Cass Ave., Lemont, IL 60439, promano@anl.gov, alund@anl.gov
\and
$^{\dagger}$Georgia Institute of Technology, 770 State St. NW, Atlanta, GA 30318, dasindrew@gatech.edu \and
$^{\ddag}$Institute of Nuclear and New Energy Technology, Tsinghua University, Beijing, China, jingang@tsinghua.edu.cn
}

\disclaimer{The submitted manuscript has been created by UChicago Argonne, LLC,
Operator of Argonne National Laboratory (``Argonne'').  Argonne, a U.S.
Department of Energy Office of Science laboratory, is operated under Contract
No. DE-AC02-06CH11357.  The U.S. Government retains for itself, and others
acting on its behalf, a paid-up nonexclusive, irrevocable worldwide license in
said article to reproduce, prepare derivative works, distribute copies to the
public, and perform publicly and display publicly, by or on behalf of the
Government. The Department of Energy will provide public access to these results
of federally sponsored research in accordance with the DOE Public Access Plan.
http://energy.gov/downloads/doe-public-access-plan.}


\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Monte Code particle transport codes are often used to calculate the distribution
of energy deposition, otherwise known as the power distribution, in nuclear
reactors. Multiplying the fission reaction rate by a constant will give a rough
estimate of the energy deposition. However, it does not account for the
redistribution of energy due to the transport of neutral secondary particles
away from a collision site. In the present work, we describe the methods used to
calculate energy deposition in OpenMC, a community-developed open source Monte
Carlo particle transport code~\cite{romano2015ane1}. With the recent inclusion
of photon transport in OpenMC~\cite{lund2018anl}, one can now obtain
very accurate energy deposition distributions with OpenMC.

%GAIL - doyou mean traverse the problem space?
As particles traverse a problem, some portion of their energy is deposited at
collision sites. The deposition occurs when charged particles, including
electrons and recoil nuclei, undergo electromagnetic interactions with
surrounding electons and ions. The total heating rate can be written as
\begin{equation}
    H(E) = \phi(E)\sum_i \rho_i \sum_r k_{i, r}(E),
\end{equation}
where $\phi(E)$ is the scalar flux at energy $E$, $\rho_i$ is the density of
nuclide $i$, and $k_{i, r}$ is the Kerma (kinetic energy released in materials)
coefficient for reaction $r$ of isotope $i$. The heating rate, $H(E)$, has units
of energy per time, typically eV/s. Kerma has units of energy $\times$
cross section (e.g., eV-barn) and can therefore be used much like a reaction
cross section for tallying energy deposition.

Kerma coefficients can be computed by using a nuclear data processing code such as
NJOY (specifically the HEATR module), which uses an energy balance method to
calculate Kerma as
\begin{equation}
    \label{eq:kerma}
    k_r(E) = \left(E + Q_r - \bar{E}_{r, \text{n}}
    - \bar{E}_{r, \gamma}\right)\sigma_{r}(E),
\end{equation}
where $Q_r$ is the $Q$ value of reaction $r$, $\bar{E}_{r,\text{n}}$ is the
average energy carried away by secondary neutrons, $\bar{E}_{r,\gamma}$ is the
average energy carried away by secondary photons, and $\sigma_r(E)$ is the
microscopic cross section for reaction $r$. Without loss of generality, we have
omitted the $i$ subscript, and it is to be understood that $k_r$ represents the
Kerma for a reaction in a specific nuclide. The term in parentheses in
\cref{eq:kerma} is known as the \emph{heating number} and represents the energy
put into the reaction (the incident neutron energy and the $Q$ value) minus the
energy transported away by neutral secondary particles.

During a fission event, there are potentially many secondary particles, and all
must be considered. The total energy released in a fission event is typically
divided into the following categories:
\begin{equation*}
\begin{split}
  \efr  &\equiv \text{energy of fission fragments} \\
  \enp  &\equiv \text{energy of prompt fission neutrons} \\
  \ened &\equiv \text{energy of delayed fission neutrons} \\
  \egp  &\equiv \text{energy of prompt fission photons} \\
  \egd  &\equiv \text{energy of delayed fission photons} \\
  \eb   &\equiv \text{energy of released $\upbeta$ particles} \\
  \enu  &\equiv \text{energy of neutrinos}
\end{split}
\end{equation*}
These components are defined in MF=1, MT=458 data in an ENDF incident neutron
sublibrary evaluation. All these quantities may depend upon incident neutron
energy; but to make the following presentation cleaner, we do not show this dependence.
NJOY assumes that the $Q$ value for fission is equal to the prompt
energy release minus the incident neutron energy:
\begin{equation}
    \label{eq:njoy-fissq}
    Q_f = \efr + \enp + \egp - E,
\end{equation}
which results in the following expression for the fission Kerma:
\begin{equation}
    k_f(E) = \left[\efr + \enp + \egp - \bar{E}_{f,\text{n}} - \bar{E}_{f,\gamma}\right]\sigma_f(E).
\end{equation}
If the secondary neutron and photon yields and energies are consistent with the
components of fission energy release in MT=458, we then have $\enp =
\bar{E}_{f,\text{n}}$ and $\egp = \bar{E}_{f,\gamma}$, which results in a
fission Kerma of
\begin{equation}
    \label{eq:njoy-kerma}
    k_f(E) = \efr \sigma_f(E).
\end{equation}

Several problems arise with the use of \cref{eq:njoy-kerma} to tally energy
deposition from fission in a Monte Carlo transport simulation. First and
foremost, it does not account for the energy deposited by photons and electrons
that are emitted during the decay of fission products. If one could 
calculate the heating due to the decay of fission products directly using the
ENDF decay sublibrary files, such an approach would be preferrable. 
This approach is complex, however, and has not been widely studied to date.
Common
practice, then, is to assume that the energy of delayed particles is released
instantaneously, which is an appropriate assumption for reactors operating at
steady state. In addition to ignoring the energy of delayed particles, when
Kerma coefficients based on \cref{eq:kerma} are used in a neutron-only
calculation, the energy from prompt photons is also ignored. In this paper, we
discuss how these two problems have been addressed in OpenMC.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}

The method for handling energy deposition in OpenMC follows the approach
introduced by Griesheimer and Stedry~\cite{griesheimer2013mc}. In this approach,
energy deposition arising from neutron interactions is calculated by using Kerma
coefficients. Energy deposition arising from photon interactions is calculated
by using an analog estimator when a coupled neutron--photon transport calculation
is carried out. In a neutron-only calculation, the energy of secondary photons
is accounted for through Kerma coefficients. We begin by discussing the
calculation and use of Kerma coefficients.

\subsection{Neutron Interactions}

As discussed above, the Kerma values calculated by NJOY are not sufficient for
accurately calculating energy deposition. Thus, the first step is to calculate
modified Kerma coefficients, which is handled through OpenMC's Python-based 
data-processing module~\cite{romano2017epjwoc}. Two sets of Kerma coefficients are
needed~\cite{trumbull2013mc} for each nuclide corresponding to the two desired
modes of operation:
\begin{enumerate}
    \item In a neutron-only calculation, Kerma coefficients should be calculated
    assuming that photons deposit their energy locally. Without this assumption,
    some other method would be needed to estimate how much energy is deposited
    from photons.
    \item In a coupled neutron--photon calculation, Kerma coefficients should be
    calculated assuming photons carry energy away from the reaction site, and
    photons should be allowed to deposit energy along their path.
\end{enumerate}
When HDF5 nuclear data files are being generated for OpenMC, NJOY is run with a
sequence that includes two HEATR steps: one that assumes photons carry energy
away and one that assumes they do not. OpenMC then stores the following modified
Kerma in the HDF5 file for the coupled neutron--photon case:
\begin{equation}
\label{eq:k}
k(E) = k_\text{NJOY}(E) - k_{f,\text{NJOY}}(E) + \left ( \efr +
        \eb \right ) \sigma_f(E),
\end{equation}
where $k_\text{NJOY}$ and $k_{f,\text{NJOY}}$ are the total (MT=301) and fission
(MT=318) KERMA coefficients obtained from the NJOY/HEATR calculation with photon
energy carried away. A modified Kerma for the neutron-only case is calculated as
\begin{equation}
    \label{eq:klocal}
    \begin{split}
    k_\text{local}(E) = \; &k_\text{local,NJOY}(E) - k_{f,\text{local,NJOY}}(E) \\
        &+ \left ( \efr + \egp + \egd + \eb \right )  \sigma_f(E),
    \end{split}
\end{equation}
where $k_\text{local,NJOY}$ and $k_{f,\text{local,NJOY}}$ are the total and
fission Kerma coefficients obtained from the NJOY/HEATR calculation assuming
photon energy is deposited locally. At runtime, OpenMC can be instructed to use
$k(E)$ for tallying energy deposition in a coupled neutron--photon calculation
and $k_\text{local}(E)$ for a neutron-only calculation. The primary benefit of
performing a coupled neutron--photon calculation is that it accurately captures
the redistribution of energy from the transport of photons. Additionally, in a
problem with vacuum boundary conditions, some of the photon energy may leak out
of the problem, which is not accounted for when using $k_\text{local}(E)$ and a
neutron-only calculation.
%GAIL - again, leak out of the problem seems odd

In a neutron-only calculation, using \cref{eq:klocal} will account for the
energy release of delayed photons and betas. In a coupled
neutron--photon calculation, however, \cref{eq:k} still does not account for delayed
photon energy release because ENDF incident neutron sublibrary files do not
include the yield or energy spectrum of delayed photons (since these depend on
the material composition). However, the average energy of delayed photons is
given in the components of fission energy release in MF=1, MT=458. To account
for the energy corresponding to the release of delayed photons, one assumes
that their energy spectrum is the same as that of prompt
photons~\cite{tuominen2019ane}. Specifically, one scales the yield of
prompt photons from fission,
\begin{equation}
    y'(E) = \frac{\egp + \egd}{\egp} y(E),
\end{equation}
where $y(E)$ is the normal prompt photon yield at energy $E$ and $y'(E)$ is a
scaled yield that accounts for the energy release of delayed photons.

In a $k$-eigenvalue calculation, an imbalance exists between energy release
and deposition related to the biasing of the fission source by $1/k_\text{eff}$,
as discussed at length by Griesheimer et
al.~\cite{griesheimer2020physor}. In order to address this imbalance, nonfission energy
deposition needs to be weighted by a normalization factor. Although not
rigorously correct, using $k_\text{eff}$ as the normalization factor provides an
approximate correction to the energy balance. When energy deposition is being
tallied in OpenMC during a $k$-eigenvalue calculation, the nonfission portion
of Kerma is weighted by the most recent estimate of $k_\text{eff}$:
\begin{equation}
    \tilde{k}(E) = \left ( k(E) - k_f(E) \right) k_\text{eff} + k_f(E),
\end{equation}
where $k_f(E)$ is calculated as $\left ( \efr + \eb \right ) \sigma_f(E)$ for a
coupled neutron--photon calculation or $\left ( \efr + \egp + \egd + \eb \right)
\sigma_f(E)$ in a neutron-only calculation. Additionally, photons born from
nonfission reactions are given a particle weight that is a factor of
$k_\text{eff}$ higher than normal.


\subsection{Photon and Electron Interactions}

Determining energy deposition from photon and electron reactions is simpler than
from neutrons because the dominant physical processes do not result in nuclear
transmutation. In principle, Kerma coefficients for photon interactions can be
calculated in the same manner as Kerma coefficients for neutron interactions.
The calculation becomes difficult, however, once one considers the tightly coupled nature of
photon--electron transport. Instead, the approach in OpenMC is to use the pre-
and postcollision energies to estimate the energy deposited. Namely, at each
collision, the energy deposited is calculated as
\begin{equation}
    w \left ( E - E' - \sum_i \mathcal{E}_i \right ),
\end{equation}
where $w$ is the weight of the particle, $E$ is the precollision energy, $E'$
is the postcollision energy (or zero if the particle does not survive the
collision), and $\mathcal{E}_i$ is the energy of the $i$th secondary particle.
This score applies whether the particle in question is a photon, electron, or
positron. Although OpenMC does not transport electrons or positrons, it does use
a thick-target bremsstrahlung treatment that results in the production of
secondary photons. This implies that energy due to electron/positron
interactions is deposited at the site where they were originally created, a
reasonable assumption given their short range in matter. That being said, we
note that the method outlined here should work equally well when electrons and
positrons are explicitly transported.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

To test the energy deposition treatment in OpenMC, we carried out a cross-code comparison
between OpenMC and Serpent 2 using two pressurized water reactor
(PWR) single assembly models from the VERA core physics benchmark progression
problems~\cite{godfrey2014casl}, namely, models 2b and 2g. These models were
chosen because results using Serpent 2 and MCNP6 have been reported
previously~\cite{tuominen2019ane}. For both models, the geometry is a single
17$\times$17 fuel assembly with 3.1\% enriched UO$_2$ fuel. The problem has
reflective boundary conditions in the $x$ and $y$ directions and extends to
positive/negative infinity in the $z$ direction. Model 2b contains 25 empty
guide tubes; 24 of the guide tubes are replaced with silver-indium-cadmium
(Ag-In-Cd) control rods in model 2g, leaving only the central guide tube.

A coupled neutron--photon simulation was run for each of the benchmark models.
The run strategy was to use 50 inactive generations, 100 active generations, and
1 million particles per generation. ENDF/B-VIII.0 interaction data was used for
both neutron and photon transport. All Serpent 2 results were taken directly
from~\cite{tuominen2019ane} rather than being calculated independently. For
these simulations, the $k_\text{eff}$ weighting factor on nonfission energy
deposition was not used to match the conditions used for the Serpent 2
simulations.\!\footnote{Serpent 2 implements a similar correction for
nonfission energy deposition; but to compare results with MCNP6, the authors
of~\cite{tuominen2019ane} had turned this feature off.}

\cref{tab:vera2b,tab:vera2g} show the percentage of energy deposited in each of
the materials in the VERA 2b and 2g models, respectively, as calculated by
Serpent 2 and OpenMC. We see that the energy deposition calculated by OpenMC
agrees with Serpent 2 within statistical uncertainty for model 2b. For model 2g,
small differences occur  outside of statistical uncertainty, but generally
the agreement is close. These differences may be caused by differences in
the physics treatments in the two codes, processing of cross sections, or model
preparation.
\begin{table}[htbp]
  \centering
  \caption{Energy deposition in VERA problem 2b.}
  \label{tab:vera2b}
  \begin{tabular}{lcc}
    \toprule
    & Serpent 2 (\%) & OpenMC (\%) \\
    \midrule
    Fuel & 96.8920 $\pm$ 0.0000 & 96.8919 $\pm$ 0.0003 \\
    Cladding & 0.9405 $\pm$ 0.0002 & 0.9404 $\pm$ 0.0001 \\
    Coolant & 2.1675 $\pm$ 0.0002 & 2.1678 $\pm$ 0.0003 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[htbp]
  \centering
  \caption{Energy deposition in VERA problem 2g.}
  \label{tab:vera2g}
  \begin{tabular}{lcc}
    \toprule
    & Serpent 2 (\%) & OpenMC (\%) \\
    \midrule
    Fuel & 94.079 $\pm$ 0.001 & 94.0723 $\pm$ 0.0008 \\
    Cladding & 1.0083 $\pm$ 0.0002 & 1.0101 $\pm$ 0.0002 \\
    Coolant & 2.6738 $\pm$ 0.0003 & 2.6763 $\pm$ 0.0004 \\
    Poison & 2.0728 $\pm$ 0.0004 & 2.0752 $\pm$ 0.0004 \\
    \bottomrule
  \end{tabular}
\end{table}

A second OpenMC simulation of VERA model 2b was carried out with the
$k_\text{eff}$ weighting factor on nonfission energy deposition applied. In
this case, OpenMC calculates an average of 199.81 $\pm$ 0.03 MeV per fission,
which agrees with the value of 199.817 $\pm$ 0.004 MeV per fission reported for
Serpent 2~\cite{tuominen2019ane}.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

A new capability to accurately calculate energy deposition has been implemented
in OpenMC. The methodology relies on calculating two sets of Kerma coefficients:
one for coupled neutron--photon transport simulations and one for neutron-only
simulations. Energy deposited in photon and electron interactions is calculated
directly by using pre- and postcollision information. Modifications to photon
yields at runtime account for delayed photon energy release and prevent an
imbalance between energy release and deposition in $k$-eigenvalue calculations.

Simulations of two PWR assembly models from the VERA core physics benchmark
progression problems were carried out, and energy deposition was tallied in each
material. For the VERA 2b model, the results using OpenMC were within
statistical uncertainty of previously reported results using Serpent 2. OpenMC
results on the VERA 2g model show small differences when compared with Serpent 2
results, but generally the agreement is close.

Energy deposition results obtained with OpenMC are only as good as the
underlying nuclear data used in a simulation. Because the energy deposited will
be influenced by integrated neutron cross sections, secondary angle and energy
distributions, photon production, photon cross sections, and fission energy
release data, bad data can cause problems in many places, as has
been documented in the
literature~\cite{trumbull2013mc,tuominen2019ane}. Even when the data is
perfect, improvements still can be made; for instance, it would be
better to use known compositions along with ENDF decay sublibrary files to
calculate energy deposition from delayed emission of neutrons, photons, and beta
particles.

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}

This research was supported by the Exascale Computing Project (17-SC-20-SC), a
collaborative effort of the U.S. Department of Energy Office of Science and the
National Nuclear Security Administration. The material was based upon work supported by the 
U.S. Department of Energy, Office of Science, under
contract DE-AC02-06CH11357.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{ans}
\bibliography{references}
%GAIL - note that ref. 7 does not state that it wes to have apeared in PHYSOR 2020 (canceled; the open access papers will be published in the proceedings on the conference website).

\end{document}

